Approximation as Differentiation: Notes

### 2025-05-09 main/formal-structures.tex
Minor corrections.

\noindent The $\partial X$ is intended to suggest the connection between the Grothendieck construction and the
idea of tangent spaces and derivatives. To understand how the $\partial X$ components compose, consider any
$(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial
Z)$ in $\Grothendieck{}T$. We have:

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
$\partial X$ in $\Fam(X,\cat{C})$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are
functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in
$\Fam(X,\cat{C})$.

@@

Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
$x \in X$, a bounded lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs
$(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x):
\partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition $(g,
\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
function $f$.

>>>>

\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
construction and the idea of tangent spaces and derivatives. To understand how the $\partial X$ components
compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in $\Fam(X,\cat{C})$.

@@

Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
$x \in X$, a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all
pairs $(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial
f(x): \partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition
$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
function $f$.
